Breakable Elastic-plastic Constraint for\\Rigid Body Simulation

### 2017-09-30 pdocs/thesis/epj-jcgt.tex
work on jcgt article

%  Simo.Nikula@gmail.com
\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse-based Physics}
\def\lut{Lappeenranta University of Technology}

@@

\begin{document}
\newcommand{\cbullet}[0]{Bullet Physics}
\title{Breakable Elastic-plastic Constraint for\\Rigid Body Simulation}

@@

\teaser{
  \includegraphics[width=5cm]{figs/demolisher-done}
  \label{fig:teaser}
  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D allows more realistic simulation of many scenarios.}
}

@@

Sample programs with source code are made available to allow developers already using \cbullet\
and Box2D to add plasticity into their simulations.
\end{abstract}

@@

\paragraph{Our approach}
This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Maximum forces and moments can be estimated based on the plastic section modulus or
by defining maximum elastic displacement.
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material and geometry based limit. The elastic part of deformation is modelled by employing 
modification of an existing constraints.

@@

Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around the $x$-axis 
and moment around the $z$-axis, respectively:

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y,
\end{equation}

@@

Maximum moment around $\vec{r}_{anc}^{\,i} $
is an integral of distance $d$ from the joint point
and shear yield stress $\tau_y$ as:

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y.
\end{equation}

@@

\begin {table}
\small
\begin{center}
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
\end{tabular}
\end{center}
\caption{Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
\label{tab:maxForces} 
\end {table}

For torque there is a closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where $b$ and $h$ are similar.
Better approximation for any given $b$ and $h$ can be obtained 
by integrating distance from the center of the joint over cross section and
multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
is shown in Listing \ref{lst:octave-mp}.

\lstset{language=octave}
\begin{lstlisting}[caption={Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}, 
label={lst:octave-mp}, float]
b=0.01; h=0.01; fy=200e6;
wpy=fy/2*dblquad(@(x,z)...
 sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
38.2
\end{lstlisting}

@@

\begin{enumerate}
\item H
\item New neutral position is calculated so that elastic part is subtracted from current position.
\item Plastic strain or rotation is incremented.
\item Current plastic strain or rotation is compared to maximum value and joint is broken if 
current value exceeds predefined maximum value. This part can be finetuned e.g. so that joint is 
not broken during compression if connected parts are made of material which can handle large
strains due to compressive forces. 
\end{enumerate}

\section{Implementation for \cbullet}

@@

\item \cbullet source with plasticity extensions 
 \href{https://github.com/simo-11/bullet3}{https://github.com/simo-11/bullet3}
\item two minute video demonstrating \cbullet implementation 
 \href{https://youtu.be/_QNbv1-f_ao}{https://youtu.be/_QNbv1-f_ao}
\item 2 MB executable used for \cbullet video {https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}{https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}
\end{enumerate}

>>>>

%  Simo.Nikula@gmail.com
\def\mytitle{Breakable Elastic-plastic Constraint for Impulse-based Physics Engines}
\def\lut{Lappeenranta University of Technology}

@@

\begin{document}
\title{Breakable Elastic-plastic Constraint for\\Rigid Body Simulation}

@@

\teaser{
  \includegraphics[width=5cm]{figs/demolisher-done}
  \label{fig:teaser}
  \caption{Adding ductile breakable constraint to popular open source physics engine Box2D allows more realistic simulation of many scenarios.}
}

@@

Sample program with source code are made available to allow developers already using 
Box2D to add plasticity into their simulations.
\end{abstract}

@@

\paragraph{Our approach}
This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Maximum forces and moments can be estimated based on the plastic section modulus or
by defining maximum elastic displacement or based on desired capacity.
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material and geometry based limit. The elastic part of deformation is modelled by employing 
modification of an existing constraints.

This paper follows idea of
allowing game designers to get desired behaviour without diving into details of structural engineering.
Idea was presented by Catto, \cite{ecsc}. Basic idea is that game designer selects spring frequency so that
integration is stable. Downside is that constraints will be quite soft. In out method game designers get new option
to use rigid-plastic constraints.

@@

Maximum moment is integral of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around  the $z$-axis:

@@

@@

\begin {table}
\small
\begin{center}
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
\end{tabular}
\end{center}
\caption{Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
\label{tab:maxForces} 
\end {table}

@@

\paragraph{Car}

@@

\end{enumerate}
